---
title: "Regression Assignment 1"
author: "Gabriel J. Odom, PhD, ThD"
date: "November 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Regression Assignment 1: Due Sunday, 11 November at 23:59PM
*For help with Rmarkdown for reports, see this [white paper](http://www.stat.cmu.edu/~cshalizi/rmarkdown/#math-in-r-markdown) from Carnegie Mellon University's Department of Statistics and Data Science.*


### 1. Create a Regression Repository on GitHub
Just like we did for the bootstrap module, create a repository on GitHub to store our work on regression. Call it `BST650_Regression` or something similar.


### 2. Primer on Partial Derivatives
Complete the following exercises. Turn in the work for 2 (the worked derivatives) to me next class on paper.

1. Work through this Khan Academy article (stop at the "A more formal definition" section): https://www.khanacademy.org/math/multivariable-calculus/multivariable-derivatives/partial-derivative-and-gradient-articles/a/introduction-to-partial-derivatives
2. Find both the partial derivatives (with respect to $\mu$ and $\sigma$) of the normal log-likelihood
\[
\log(L(\mu, \sigma | \textbf{x})) = \sum_{i = 1}^n \log\left( \frac{1}{\sqrt{2\pi\sigma^2}} \exp\left[ -\frac{(x_i - \mu)^2}{2\sigma^2} \right] \right).
\]
Set these partial derivatives equal to zero, and solve for $\mu$ and $\sigma$. Do you recognize these formulas?


### 3. Calculate Univariate Regression Estimates
Consider on the simple linear regression model we covered in class:
```{r regression_example}
wcgs_df <- read.csv("../data/wcgs.csv")

plot(
  x = wcgs_df$dbp,
  y = wcgs_df$sbp,
  xlab = "Diastolic Blood Pressure (mmHg)",
  ylab = "Systolic Blood Pressure (mmHg)"
)
lm(sbp ~ dbp, data = wcgs_df)
```

Based on this model output, the slope ($\beta_1$) is approximately 1.2, while the intercept ($\beta_0$) is about 30.1. 

Write a function which takes in any two numeric vectors (`x` and `y`) and returns a named vector of the slope and intercept of the linear model $\hat{\textbf{y}} = \beta_0 + \beta_1\textbf{x}$. Your function should use formulas (from your other classes) to calculate these estimates directly (that is, without a call to the `lm()` or any other modelling function in R).

**If you don't have enough time to complete part 3, that's ok. But at least start on it.**