---
title: "Plotting Distributions"
author: "Gabriel Odom"
date: "8/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Random variables in `R` are shown on https://stat.ethz.ch/R-manual/R-devel/library/stats/html/Distributions.html

There are two functions we will use: `hist()` and `plot(density())`. There is another function, `summary()` which will give us the mean and 5-number summary of a numeric vector.


# Continuous Random Variables

## The Normal Distribution

### Sample Size of 10000
```{r}
# Draw Random Sample: sqrt(5) ~= 2.24
xFullN <- rnorm(10000, mean = 2, sd = sqrt(5))

# Histogram
hist(xFullN)

# Summary
summary(xFullN)
var(xFullN)

# Density Plots
density(xFullN)
plot(density(xFullN))

# Scatterplots
plot(xFullN)
```

### Sample Size of 5
```{r}

# Histogram
hist(xFullN[1:5])

# Summary
summary(xFullN[1:5])
var(xFullN[1:5])

# Density Plots
density(xFullN[1:5])
plot(density(xFullN[1:5]))
```

### Sample Size of 15
```{r}
# Histogram
hist(xFullN[1:15])

# Summary
summary(xFullN[1:15])
var(xFullN[1:15])

# Density Plots
plot(density(xFullN[1:15]))
```

### Sample Size of 25
```{r}
# Histogram
hist(xFullN[1:25])

# Summary
summary(xFullN[1:25])
var(xFullN[1:25])

# Density Plots
plot(density(xFullN[1:25]))
```

### Sample Size of 50
```{r}
# Histogram
hist(xFullN[1:50])

# Summary
summary(xFullN[1:50])
var(xFullN[1:50])

# Density Plots
plot(density(xFullN[1:50]))
```


## The Student's $t$ Distribution

### Sample Size of 10000
```{r}
# Draw Random Sample: sqrt(5) ~= 2.24
xFullSt <- rt(n = 10000, df = 4)

# Histogram
hist(xFullSt)

# Summary
summary(xFullSt)
var(xFullSt)

# Density Plots
plot(density(xFullSt))

# Scatterplot
plot(xFullSt)
```


## The $\chi ^ 2$ Distribution


## The $F$ Distribution



# Discrete Distributions

## The Binomial Distribution


## The Negative Binomial Distribution


## The Poisson Distribution

