---
title: "Statistical Distributions"
author: "Gabriel Odom"
date: "8/30/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Random variables in `R` are shown on https://stat.ethz.ch/R-manual/R-devel/library/stats/html/Distributions.html

# Discrete Random Variables

## Distributions
We have a few common examples of discrete random variables. Each distribution also has a corresponding implementation in `R`.
1. The Bernoulli Distribution: <https://en.wikipedia.org/wiki/Bernoulli_distribution>
```{r}
# Random Bernoulli with a fair coin
rbinom(n = 1, size = 1, prob = 0.5)

# Weighted coin
rbinom(n = 1, size = 1, prob = 0.75)
```

2. Binomial Distribution: <https://en.wikipedia.org/wiki/Binomial_distribution>
```{r}
# Flip 10 fair coins and record the results
rbinom(n = 10, size = 1, prob = 0.5)

# Flip 10 fair coins and record the number of successes
rbinom(n = 1, size = 10, prob = 0.5)
```

3. Poisson Distribution: <https://en.wikipedia.org/wiki/Poisson_distribution>
```{r}
# How many horse kicks in each group of 10 soldiers? We expect 2 kicks
rpois(n = 10, lambda = 2)

# Now the horses are angry...
rpois(n = 10, lambda = 4)
```

## Moments of these Distributions
The common moments we would care about are the *first* and *second* moments, from which we calculate the mean and variance, respectively.

### Binomial
```{r}
# Random draw of five fair coins
xBinom <- rbinom(n = 5, size = 1, prob = 0.5)
xBinom

# mean
mean(xBinom)

# variance
var(xBinom)


# Random draw of 100 fair coins
xBinom2 <- rbinom(n = 100, size = 1, prob = 0.5)

# mean
mean(xBinom2)

# variance
var(xBinom2)
```


# Continuous Random Variables