---
title: "Checking Regression Assumptions"
author: "Gabriel Odom"
date: "11/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
We have built a regression model assessing the relationship between blood cholesterol and age, weight, and systolic blood pressure.

What are the five assumptions of multiple linear regression?

0. The linear model is appropriate
1. Errors are normally distributed
2. Errors are centred at 0
3. Errors have constant variance (Homoskedasticity)
4. Errors are independent
5. No multicollinearity

We can encapsulate the first four into a single equation:
\[
\varepsilon_i\ \sim \text{i.i.d.}\ N(0, \sigma)
\]



## Setup
We load the packages and data we need:
```{r}
library(tidyverse)
wcgs_df <- read_csv("../data/wcgs.csv")
```

Now we can construct the linear model in order to test the assumptions above.
```{r}
chol_mod <- lm(chol ~ age + weight + sbp, data = wcgs_df)
```



## Assumptions of Multiple Linear Regression

### Assumption 1: Model Errors are Normally Distributed
Overall, the residuals appear to be normally distributed, based on the histogram, but we may have a few outliers. When we inspect the QQ-plot, we probably have an outlier.
```{r}
str(chol_mod)
# Histogram
hist(chol_mod$residuals)

# QQ-Plot
qqnorm(chol_mod$residuals)
qqline(chol_mod$residuals)
```

Try a transformation:
```{r}
hist(wcgs_df$chol)
cholSq_mod <- lm(sqrt(chol) ~ age + weight + sbp, data = wcgs_df)
hist(cholSq_mod$residuals)

qqnorm(cholSq_mod$residuals)
qqline(cholSq_mod$residuals)
```

That's probably not any better. Let's leave the data alone...


### Assumption 2: Model Errors are Centred at 0
The quick and dirty way would be to look at the centre of the histogram. We can also perform a statistical test:
```{r}
t.test(chol_mod$residuals, alternative = "two.sided", mu = 0)
```

0 is dead-set in the middle of the confidence interval. We're rather confident that the mean of the residuals is close enough to 0.


### Assumption 3: Model Errors are Homoskedastic


### Assumption 4: Model Errors are Independent


### Assumption 5: Model Predictors are not Perfectly Correlated